---
title: "Asssignment 3, Task 3"
author: "Evie Vermeer"
date: "10/24/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
options(digits=3)
library(here)
library(tidyverse)
library(ggplot2)
library(janitor)
library(kableExtra)
library(lubridate)
```

```{r}
# Read in data and clean 
urchins <- read_csv(here("data","urchins.csv")) %>% clean_names()
```

### Section 1: Size comparisons, purple and red urchins
```{r}
# Data vizualization
ggplot(data = urchins, aes(x=size,fill=common_name)) + #using urchin data, puting size on x-axis and assigning colors by urchin type
  geom_histogram(bins = 10, position = "identity")+ # adjusting no. of bins 
  facet_wrap(~common_name)+ # facet wrapping to create a histogram for each urchin type
  labs(x = "Urchin Diameter (mm)", 
       y= "Number of Urchins",
       title = "Size Distribution Comparisons, Purple and Red Urchins",
       fill ="Urchin Type") + # adding axis labels, along with titles for histograms and legends
  scale_fill_manual(values = c("Purple Urchin"="blueviolet","Red Urchin"="firebrick3")) # customized colors for each histogram
```

```{r}
#Descriptive statistics table
urchins_stats <- urchins %>% #creating new data set
  group_by(common_name) %>% #grouping by common name 
  summarize(mean_size = mean(size, na.rm = TRUE),
            sd_size = sd(size, na.rm = TRUE),
            sample_size = n()) #creating varialbes for mean, sd and sample size for each type of urchin 

urchins_stats %>% 
  kable(col.names = c("Species",
                      "Mean (mm)",
                      "Standard Deviation (mm)",
                      "Sample Size (n)"), #adding variables to table
                    caption = "Descriptive Statistics of Urchin Diameters by Species", 
                    digits = 3) %>% # adding caption and significant figure parameters
  kable_styling(full_width = FALSE) # formatting
```

##### These data show that purple urchins are generally smaller in diameter than red urchins. Specifically, the average diameter of a purple urchin is 34% less than the average diameter of a red urchin (4.47 and 6.78 mm, respectively). The histogram above shows the the most common size of a purple urchin is less than the most common size of a red urchin. 

### Section 2: Purple Urchin Counts by site and year

```{r}
purple_urchins <- urchins %>% # creating new data set for purple urchins 
  filter(common_name=="Purple Urchin") %>% #only common names for purple urchin 
  mutate(date_new = mdy(date)) %>% 
  mutate(year = year(date_new)) %>% # changing date variable to new ones, ultimately just isolating year
  mutate(full_site = case_when(
          site %in% c("NAPL") ~ "Naples Reef",
          site %in% c("AQUE") ~ "Arroyo Quemado Reef",
          site %in% c("CARP") ~ "Carpinteria Reef",
          site %in% c("MOHK") ~ "Mohawk Reef",
          site %in% c("IVEE") ~ "Isla Vista Reef")) # adding full names of reefs

```

```{r}
purple_urchins_stats <- purple_urchins %>% 
  group_by(full_site,year) %>% # grouping data by reefs
  count(common_name) %>% #counting no. of purple urchins by each reef
  na.omit()
```

```{r}
ggplot(data = purple_urchins_stats, aes(x=year, y=n,fill=full_site,color=full_site),)+ 
  # creating plot with new dataset, adding axis criteria, legend and assigning colors 
  geom_line(stat = "identity")+ # making line graph
  theme_bw()+ # theme
  theme(legend.title = element_blank())+ #removing legend title (self-explanatory)
  labs(x="Year",
       y="Number of Purple Urchins",
       title="Number of Purple Urchins at Each Site by Year")+ #adding axis labels and title  
  scale_x_continuous(breaks = seq(2008,2020,2)) #adjusting scale of x-axis
 
# Can also use scale_x_discrete(limit = c(2008-2020))
```

##### The data shown lead me to a few preliminary observations. At first glance, it would seem to show that purple sea urchins are declining at these sites year over year. However, we would need to aggregate the number of urchins at each site for this timeframe and determine some summary statistics such as the mean and the data spread, and then compare these data and their progressions in that context. A regression of these trends would be helpful as well.  


