---
title: "Assignment 3, Task 4"
author: "Evie Vermeer"
date: "10/24/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
options(digits=3)
library(here)
library(tidyverse)
library(ggplot2)
library(janitor)
library(kableExtra)
library(lubridate) 
library(plotly)

```

```{r}
# Read in data and clean 
urchins <- read_csv(here("data","urchins.csv")) %>% clean_names()
```

```{r}
mean_size_purple_urchins <- urchins %>% # creating new dataset
  filter(common_name=="Purple Urchin") %>% #filtering by purple urchin
  group_by(treatment) %>% #grouping by treatment
  summarize(mean_size=mean(size)) #adding variable for mean of size
```


```{r}
naples_purple_urchins_2010 <- urchins %>% # Creating new dataset
  mutate(date_new = ymd(date)) %>% 
  mutate(year = year(date_new)) %>% # New variable of year after mutating old variable/isolating year
  filter(year == "2010",common_name == "Purple Urchin", site == "NAPL") # filtering by year, species and reef
  

```

```{r}
ggplot(data=naples_purple_urchins_2010, aes(x=size,fill=treatment))+ # new plot with x-axis:size and using variable: treatment
      geom_histogram(bins=8, alpha=.4,  position ="identity")+ # adjusting bins (and opacity for fun)
  facet_wrap(~treatment)+ # facet wrapping to get two histograms for each treatment
  labs(x= "Size (cm)",
           y= "Number of Urchins",
           title="Purple Urchin Sizes on Naples Reef - 2010",
           subtitle= "Comparison of Annual (Kelp Removal) vs. Control (No Kelp Removal)",
           fill="Treatment") # adding axis labels, and titles(/subtitles)
  
```

```{r}
ggplot(data=naples_purple_urchins_2010, aes(sample=size, color =treatment))+ # plotting with same dataset, assigning color by treatment
      geom_qq(shape = 4, size = 4)+ #creating quartile-quartile plot
      geom_qq_line(color="black")+ #adding quartile line
      facet_wrap(~treatment) + # facet wrapping to get two plots for each treatment
      labs(x= "Offset from Median (cm)",
           y= "Size (cm)",
               title="Quantile Quantile Plot: Annual (Kelp Removal) vs. Control (No Kelp Removal)") #adding axis labels and title
```
 
### As the QQ plot and viewing of the histograms above reveal, the data for the annual treatment is not normally distributed while the data for the control treatment mostly is. While the data for the kelp removal treatment do not have a normal distribution, their mean and mode are both slightly lower than the data for the control (no kelp removal). This is interesting to reflect upon because both sample sizes are (fairly) robust.

```{r}
naples_purple_urchins_2010_stats <- naples_purple_urchins_2010 %>% # creating dataset for statistics
  group_by(treatment) %>% #grouping by each treatment
  summarize(mean_size=mean(size, na.rm= TRUE), 
            sd_size=sd(size, na.rm= TRUE),
            sample_size=n()) #creating variables for mean, sd and sample size

naples_purple_urchins_2010_stats %>% 
  kable(col.names = c("Treatment",
                      "Mean (mm)",
                      "Standard Deviation(mm)",
                      "Sample size (n)"), #creating table with data, adding headers
                    caption = "Descriptive Statistics for Purple Urchins on Naples Reef - 2010") %>% #adding title
  kable_styling(full_width=FALSE) #formatting

# Can also use the round() function to adjust significant figures 
```



